---
title: "802 Project Summary"
authors: 
  Maksuda Aktar Toma,
  Jo Charbonneau,
  Ryan Lalicker
date: today
date-format: long
execute: 
  echo: false
  warning: false
columns: 2
format:
  pdf: 
    fig-align: center
    fig-width: 6
    fig-height: 4
editor: 
  markdown: 
    wrap: sentence
---

```{r,  fig.pos="H"}
#| label: data-setup
#| echo: false
#| eval: true

# Put most/all libraries in this code chunk and read in needed data files.

library(knitr)
library(dplyr)
library(ggplot2)
library(naniar)
library(reshape2)
library(GGally)
library(janitor)
library(emmeans)
library(MASS)
library(multcomp)
library(lme4)
library(nnet)

data <- read.csv("rptm_simulation.csv")
head(data)


```

```{r,  fig.pos="H"}
#| label: Scratch-work-1
#| echo: false
#| eval: false

```

# Introduction


For more information on the experiment, the data, or any other files used in this paper see our [Github page](https://github.com/maksudatoma/Stat-802-Project) which can be found at https://github.com/maksudatoma/Stat-802-Project.
The coding languages used in the paper are R and SAS.
The corresponding code can be found in *Appendix A - R Code* and *Appendix B - SAS Code* respectively.

## Variables


## Missing Values



```{r,,fig.pos="H"}
#| label: fig-some-label
#| echo: false
#| eval: true
#| fig-cap: ""

# To create a plot copy this code segment and change the name of the label above.  You can then enter the label in fig-cap.  

```


## Summary Statistics



```{r,  fig.pos="H"}
#| label: fig-two-plots
#| echo: false
#| eval: true
#| fig-cap: "Two plots"
#| layout-ncol: 2
#| fig-subcap: 
#|  - "First"
#|  - "Second"
#| fig-width: 6
#| fig-height: 4

# This allows you to make two plots side by side.  If you want to plot 4 plots you can use `#| layout-nrow: 2` to get a 2x2 setup.

```


## Exploring the Data




# Model


Models can be written as 

$$
Y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha \beta)_{ij}+u_k+e_{ijk}
$$ 

where $Y_{ijk}$ represents the final weight for the $ith$ treatment, the $j$th sex, and the $k$th sire.
$\mu$ represents the overall mean, while $\alpha_i$ is the fixed effect for the $ith$ treatment, $\beta_j$ is the fixed effect for the $j$th sex, and $(\alpha \beta)_{ij}$ is their interaction.
The random sire effect is represented by $u_k$ and we assume $u_k~N(0,\sigma_k^2)$.
We also assume the residual term $e_{ijkl}$ is distributed as $~N(0,\sigma^2)$.


# Conclusion 



# Future Work



\newpage

# References

::: {#refs}
:::

\newpage

# Appendix A - R Code

```{r,  fig.pos="H"}
#| label: appendix A
#| echo: true
#| eval: false

```

\newpage

# Appendix B - SAS Code

``` sas

```